---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
# ppmnet

<!-- badges: start -->
<!-- badges: end -->

`ppmnet` is an R package for fitting regularized spatial point process models.

Models are fit via penalized likelihood for Poisson point processes, and via penalized composite likelihood for Gibbs point processes. 
The model-fitting procedure is carried out by `glmnet`, which implements the lasso and elastic net penalties.
A number of methods are provided for plotting, prediction, validation, and model selection on the basis of (composite) information criteria.


## Installation

You can install `ppmnet` from [GitHub](https://github.com/) with:

```{r installation, eval = FALSE}
devtools::install_github("jeffdaniel/ppmnet")
library(ppmnet)
```

## Example

Fit a regularized inhomogeneous Poisson point process model to the `bei` dataset
included with `spatstat`.
```{r beifit}
fit <- ppmnet(quadscheme(bei), bei.extra)
```

Select the optimal model from the regularization path on the basis of minimum AIC and plot the model's predicted intensity surface.
```{r beiplot}
lambda <- select(fit, "AIC")$lambda
plot(predict(fit, bei.extra, s = lambda))
```

Examine the smoothed residuals for the selected model.
```{r resplot}
plot(Smooth(residuals(fit, s = lambda)))
```


Fit an inhomogeneous area-interaction model to the `bei` dataset via penalized logistic composite likelihood.
```{r aifit}
fit.ai <- ppmnet(quadscheme.logi(bei), bei.extra,
                 interaction = AreaInter(10), method = "logi")
```

## Citation

Daniel, J., Horrocks, J., and Umphrey, G. J. Penalized composite likelihoods for inhomogeneous Gibbs point process models. *Computational Statistics and Data Analysis*, **124**, 104--116. https://doi.org/10.1016/j.csda.2018.02.005
